---
title: "Heatmap by Day of Week and Hour of Day"
---

This example pulls **visits** by **device type** by _hour_ for the last 30 days and then plots the data as a heatmap. In this example, we simply use `ggplot` functionality. Another approach is to use the `highcharter` package as described in [this example](http://www.dartistics.com/example5.html).

We're actually going to plot the _average_ visits by hour-weekday rather than the total. Otherwise, we'd be stacking the deck for a couple of weekdays by pulling 30 days rather than a divisible-by-7 number of days.

# Setup/Config

```{r child='includes/z_setup.Rmd'}
```

# Pull the Data

This is a fairly simple query. We're going to need both the date (to determine day of week) and the hour of the day. When we specify a `date.granularity` of `hour`, we actually get _both_ a `datetime` column _and_ an `hour` column returned, which actually simplifies our data munging a bit.

For giggles, we're also going to pull the data broken out by device type so that we can do a little extra visualization. This means we'll first have to roll up the data to _remove_ the device type for our initial visualization, but that's a simple enough step. 

```{r data-pull, message=FALSE, warning=FALSE, results='hide'}
# Pull the data. See ?QueueTrended() for additional parameters.
aa_data <- QueueTrended(rsid,
                        date.from = start_date,
                        date.to = end_date,
                        metrics = "visits",
                        elements = "mobiledevicetype",
                        date.granularity = "hour")
  
# Go ahead and do a quick inspection of the data that was returned. This isn't required,
# but it's a good check along the way.
head(aa_data)
```

```{r data-pull-display, echo=FALSE, message=FALSE, warning=FALSE}
# Nicer display of the head
kable(head(aa_data))
```

# Data Munging

We need to do a few things in this step:

* Change the date to be `Date` instead of `POSIXlt` to play nice with the plotting.
* Add a column that specifies the weekday. We'll actually make this column a factor (rather than just a string) so that we can control the order that the weekdays are arranged (otherwise, we would wind up with the weekdays in alphabetical order, which would be unhelpful).
* Change "Other" to be "Desktop/Notebook."

```{r data-munging, message=FALSE, warning=FALSE, results='hide'}

# Convert the datetime from POSIXlt to Date. This makes the next step smoother.
aa_data$datetime <- as.Date(aa_data$datetime)


# Add a column with the weekday and then abbreviate the weekday to be just
# the first 3 letters. Then, make that column a factor with the 
# values in a specific order (the normal order for days of the week)
aa_data <- aa_data %>% 
  mutate(weekday = weekdays(datetime)) %>% 
  mutate(weekday = factor(weekday,
                          levels =  c("Sunday", "Monday", "Tuesday", "Wednesday",
                                      "Thursday", "Friday", "Saturday")))

# Clean up the table a bit by convering "Other" to "Desktop/Tablet" and
# then grabbing just the columns of interest.
aa_data <- aa_data %>% 
  mutate(Device = ifelse(name=="Other", "Desktop/Tablet", name)) %>% 
  select(datetime, weekday, hour, Device, visits) 

# Check what it looks like now
head(aa_data)
```

```{r data-munging-display, echo=FALSE, message=FALSE, warning=FALSE}
# Nicer display of the head
kable(head(aa_data))
```

Now, we're going to make two sets of data from `aa_data`:

* `aa_data_devicetype` -- this will be a data frame for which we get the average (mean) visits for each weekday-hour-devicetype combination
* `aa_data_total` -- this will be a data frame for which we get the average visits for just each weekday-hour combination.

```{r data-munging-2, message=FALSE, warning=FALSE, results='hide'}

# Group by the weekday, hour, and devicetype, and then summarise by getting the
# mean of the visits.
aa_data_devicetype <- aa_data %>% 
  group_by(weekday, hour, Device) %>% 
  summarise(visits = mean(visits))

# Do the same thing, but leave out devicetype
aa_data_totals <- aa_data %>% 
  group_by(weekday, hour) %>% 
  summarise(visits = mean(visits))

# Check what the second one looks like
head(aa_data_totals)

```

```{r data-munging-display-2, echo=FALSE, message=FALSE, warning=FALSE}
# Nicer display of the head
kable(head(aa_data_totals))
```

# Data Visualization

This won't be the prettiest possible plot, but we'll do a bit of tweaking to make it "better than the defaults."" If you want to read up more on ggplot, [this page on dartistics.com](http://www.dartistics.com/ggplot2.html) is worth checking out.

```{r visualize, message=FALSE, warning=FALSE}

# Create a theme that is a bit more friendly. This isn't required to do the visualization,
# but it's a painful heatmap to look at otherwise. We start with a minimalist theme... and
# then basically make it even more minimalist. Thank you, Tufte and Few.
theme_heatmap <- theme_light() +                 # Start with a minimalist theme
  theme(panel.grid = element_blank(),            # Remove the gridlines
        panel.border = element_blank(),          # Remove the border around the heatmap
        plot.title = element_text(face = "bold", # Make the title bold
                                  size = 11,     # Adjust the title size
                                  hjust = 0.5),  # Center the title
        axis.ticks = element_blank(),            # Remove the axis tickmarks
        axis.title.x = element_blank(),          # Turn off the x-axis title 
        axis.title.y = element_text(size=10),    # Adjust the size of the y-axis title
        axis.text.y = element_text(size = 8),    # Adjust the size of the y-axis labels
        axis.text.x = element_text(size = 10),   # Adjust the size of the x-axis labels
        legend.position = "none")                # Turn off the legend

# Create the plot.
gg <- ggplot(aa_data_totals, mapping = aes(x = weekday, y = hour, 
                                           fill = visits)) +
  geom_tile(colour="white") +  # This makes the heatmap (the colour is the lines in the grid)
  scale_fill_gradient(low = "#f4fff6", high="#0bb730") +  # The colour range to use
  scale_y_reverse(breaks=c(23:0), labels=c(23:0),    # Put the 0 at the top of the y-axis
                  expand = c(0,0)) +                 # Remove padding around the heatmap
  scale_x_discrete(expand = c(0,0), position = "top") +
  labs(title = "Average Visits by Day of Week / Hour of Day", y = "Hour of Day") +
  theme_heatmap  # Apply the theme defined earlier for styling

# Output the plot. You *could* just remove the "gg <-" in the code above, but it's
# generally a best practice to create a plot object and then output it, rather than
# outputting it on the fly.
print(gg)
```

If we wanted to see the actual average visits within the cells, we can take the plot object (gg) we created already and just extend it a little bit by adding `geom_text()` to it.

```{r visualize-2, message=FALSE, warning=FALSE}

# Make a new plot object that adds text in the cells
gg_labeled <- gg +
  geom_text(aes(label = sprintf("%.1f", visits)),  # sprintf formats to 2 decimals
            size = 3)                              # And we need to tweak the font size

# Output that object
print(gg_labeled)

```

Alternatively, we can quickly make three heatmaps as "facets" -- one for each device type. We'll use the `aa_data_devicetype` for this, but, other than adding a `facet_grid` component, it's pretty much the same as the original heatmap (and we can use the theme we already defined). Notice how the shading is based on _all_ of the plots. Since there is very little traffic from tablets, there is no real "peak" time relative to all traffic.

```{r visualize-3, message=FALSE, warning=FALSE, fig.height = 12}

# Create the plot.
gg_devicetype <- ggplot(aa_data_devicetype, mapping = aes(x = weekday, y = hour, 
                                                          fill = visits)) +
  geom_tile(colour="white") +  
  scale_fill_gradient(low = "#f4fff6", high="#0bb730") + 
  scale_y_reverse(breaks=c(23:0), labels=c(23:0),
                  expand = c(0,0)) +
  scale_x_discrete(expand = c(0,0), position = "top") +
  labs(title = "Average Visits by Day of Week / Hour of Day", y = "Hour of Day") +
  facet_grid(Device ~ .) +    # This is the sole addition to the plot
  theme_heatmap 

# Output the plot.
print(gg_devicetype)

```