---
title: "Apply Segment by Segment ID"
---

This example pulls the top 10 pages for the last thirty days, _for visits where the entry page included the word "Intelligence"_.  We'll do this by referencing a segment that was built in the Adobe Analytics web UI.

# Setup/Config

```{r child='includes/z_setup.Rmd'}
```

# Pull the Data

The key step here is figuring out the ID of the segment that you want to use. The easiest way to find this is to use the `GetSegments()` function:

```segs <- GetSegments(rsid)```

That will create a data frame that you can then view/filter to find the specific ID you want to use, which you can then copy and paste into the actual data pulling code.

```{r data-pull, message=FALSE, warning=FALSE, results='hide'}

# Pull the data. See ?QueueRankes() for details on the arguments available.
aa_data <- QueueRanked(rsid,
                       date.from = start_date,
                       date.to = end_date,
                       metrics = "pageviews",
                       elements = "page",
                       top = 10,
                       segment.id = "s300003965_5aad9f5c22fa9942088629ea")
  
# Go ahead and do a quick inspection of the data that was returned. This isn't required,
# but it's a good check along the way.
head(aa_data)

```

```{r data-pull-display, echo=FALSE, message=FALSE, warning=FALSE}
# Nicer display of the head
kable(head(aa_data))
```

# Data Munging

In order to keep the order in the bar chart, we need to convert the page column to be a factor. We'll reverse the order so that, when displayed in a bar chart, they'll be in descending order.

```{r data-munging, message=FALSE, warning=FALSE}
# Convert page to be a factor
aa_data$name <- factor(aa_data$name,
                       levels = rev(aa_data$name))
```

# Data Visualization

This won't be the prettiest bar chart, but let's make a horizontal bar chart with the data. Remember, in **ggplot2**, a horizontal bar chart is just a normal bar chart with `coord_flip()`.

```{r visualize, message=FALSE, warning=FALSE}

# Create the plot. Note the stat="identity"" (because the data is already aggregated) and
# the coord_flip(). And, I just can't stand it... added on the additional theme stuff to
# clean up the plot a bit more.
gg <- ggplot(aa_data, mapping = aes(x = name, y = pageviews)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  theme_light() +
  theme(panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.border = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks.y = element_blank())

# Output the plot. You *could* just remove the "gg <-" in the code above, but it's
# generally a best practice to create a plot object and then output it, rather than
# outputting it on the fly.
gg

```
