---
title: "Simple Inline Segment"
---

"Simple" may be a bit of a misnomer here. This is a "simple-in-what-we're-trying-to-do" example if not a "simple-to-actually-implement" one.

**Note**: _This example uses the `unbox()` function from the `jsonlite` package. If you do not already have `jsonlite` installed (or if you're not sure), run `install.packages("jsonlite")` in the console before running this example._

This example pulls the top 10 pages for the last thirty days, _for visits where the entry page included the word "excel"_.*  We'll do this by building a segment _dynamically_ rather than referencing a segment ID. Where this can come in handy is if you have a script where you want to work through a range of small little tweaks to one segment and re-pull the data. You don't want to build each segment in the web interface and then hardcode all those IDs! We may add an example for doing that later, but we're doing to keep this very simple for now.

This returns the exact same results as these [Apply Segment by Segment ID](segment-by-id.html) example. It just doesn't require any work in the Adobe Analytics web interface to return the results.

# Setup/Config

```{r child='includes/z_setup.Rmd'}
```

# Pull the Data

This gets built up in what can feel a bit cumbersome. _Essentially_ (but not necessarily _intuitively_), Adobe Analytics segments can have containers within containers. R has an analog: lists. A list can have a list as an element. So, we use lists to define inline segments. 

```{r data-pull, message=FALSE, warnings=FALSE, results='hide'}

# Create the actual segment object that we're going to use in the query. See ?segment_ga4()
# for details.
my_segment <- list(container = list(type = unbox("visits"),
                                    rules = data.frame(name = "Entry Page Includes 'excel'",
                                                       element = "entrypage",
                                                       operator = "contains",
                                                       value = "excel")))

# Pull the data. See ?QueueRankes() for details on the arguments available. Note that
# we're using segment.inline as an argument here.
aa_data <- QueueRanked(rsid,
                       date.from = start_date,
                       date.to = end_date,
                       metrics = "pageviews",
                       elements = "page",
                       top = 10,
                       segment.inline = my_segment)
  
# Go ahead and do a quick inspection of the data that was returned. This isn't required,
# but it's a good check along the way.
head(aa_data)
```

```{r data-pull-display, echo=FALSE, message=FALSE, warnings=FALSE}
# Nicer display of the head
kable(head(aa_data))
```

# Data Munging

In order to keep the order in the bar chart, we need to convert the page column to be a factor. We'll reverse the order so that, when displayed in a bar chart, they'll be in descending order.

```{r data-munging, message=FALSE, warning=FALSE}
# Convert page to be a factor
aa_data$name <- factor(aa_data$name,
                       levels = rev(aa_data$name))
```

# Data Visualization

This won't be the prettiest bar chart, but let's make a horizontal bar chart with the data. Remember, in **ggplot2**, a horizontal bar chart is just a normal bar chart with `coord_flip()`.

```{r visualize, message=FALSE, warning=FALSE}

# Create the plot. Note the stat="identity"" (because the data is already aggregated) and
# the coord_flip(). And, I just can't stand it... added on the additional theme stuff to
# clean up the plot a bit more.
gg <- ggplot(aa_data, mapping = aes(x = name, y = pageviews)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  theme_light() +
  theme(panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.border = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks.y = element_blank())

# Output the plot. You *could* just remove the "gg <-" in the code above, but it's
# generally a best practice to create a plot object and then output it, rather than
# outputting it on the fly.
gg

```

_* Yes. This is intentionally ironic._