---
title: "Traffic by Medium"
---

This example pulls **visits** by **device type** for the last 30 days and then plots the data on a line chart.

# Setup/Config

```{r child='includes/z_setup.Rmd'}
```

# Pull the Data

This is a simple query. 

```{r data-pull, message=FALSE, warning=FALSE, results='hide'}
# Pull the data. See ?QueueTrended() for additional parameters.
aa_data <- QueueTrended(rsid,
                        date.from = start_date,
                        date.to = end_date,
                        metrics = "visits",
                        elements = "mobiledevicetype",
                        date.granularity = "day")
  
# Go ahead and do a quick inspection of the data that was returned. This isn't required,
# but it's a good check along the way.
head(aa_data)
```

```{r data-pull-display, echo=FALSE, message=FALSE, warning=FALSE}
# Nicer display of the head
kable(head(aa_data))
```

# Data Munging

There's not a ton of data munging to be done here. We'll change the date to be `Date` instead of `POSIXlt` to play nice with the plotting, and we'll grab just the columns we need to do the data visualization. And, for giggles, we'll change "Other" to be "Desktop/Notebook."

```{r data-munging, message=FALSE, warning=FALSE, results='hide'}

# Convert the datetime from POSIXlt to Date
aa_data$datetime <- as.Date(aa_data$datetime)

# Clean up the table a bit
aa_data <- aa_data %>% 
  mutate(device = ifelse(name=="Other", "Desktop/Tablet", name)) %>% 
  select(datetime, device, visits) 

# Check what it looks like now
head(aa_data)
```

```{r data-munging-display, echo=FALSE, message=FALSE, warning=FALSE}
# Nicer display of the head
kable(head(aa_data))
```

# Data Visualization

This won't be the prettiest plot, but this example isn't diving into the details of ggplot2. If you want to read up on that, [this page on dartistics.com](http://www.dartistics.com/ggplot2.html) is worth checking out.

```{r visualize, message=FALSE, warning=FALSE}

# Create the plot.
gg <- ggplot(aa_data, mapping = aes(x = datetime, y = visits, colour = device)) +
  geom_line() +
  theme_light()

# Output the plot. You *could* just remove the "gg <-" in the code above, but it's
# generally a best practice to create a plot object and then output it, rather than
# outputting it on the fly.
gg

```